---
title: 
  "An introduction to open-source R software for wearable device data processing and analysis"
author: "Brian C. Helsel, PhD"
format: 
  pptx:
    reference-doc: "template.pptx"
output:
  pptx_document
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Set a default CRAN mirror
options(repos = "https://cran.r-project.org")

# Load necessary libraries
library(quarto)
library(ggplot2)
library(officer)
```

## Introduction

::: {.incremental}
- R is a functional programming language for statistical computing and graphics
- RStudio is an integrated development environment for R
- There are >20,000 available packages on CRAN extending the utility of R
:::

::: {.notes}
A functional programming language means that you can extend the functionality of the programming language by writing functions designed to accomplish specific tasks.

RStudio is not the only way to interact with R on your computer. For example, some users prefer to interact with R in a separate R console or integrated development environment like VS Code. R can also be run from your terminal.

Common functions are often assembled into a package for repeated use.
 
:::

## Why R?

:::{.incremental}
- Many open-source physical activity packages available on CRAN or GitHub
- Open-source code free to view on GitHub
- Reticulate is a Python-R interface that makes Python packages available to R users
- Customized scripts or packages can meet the needs of a research study. 
:::

::: {.notes}

Commonly used R packages include read.gt3x for reading binary data, PhysicalActivity for wear time processing using the Choi algorithm, and GGIR for analyzing raw acceleration sleep and physical activity data. The compositions package is not a physical activity-specific package, but is being used more by researchers now that it's being suggested that physical activity data be analyzed in a composition. Additionally, I've been working on developing a few R packages called agcounts and iFitbit. agcounts allows users to calculate ActiGraph counts from raw acceleration data and iFitbit interacts with the Fitbit API to extract data from research participants.

Working with time series data is complex, but open source code, scripts, and packages can make it easier and help you learn!

:::

## Objectives

The purpose of today's presentation is to:

::: {.incremental}
1. Introduce you to R packages used in physical activity research
2. Provide examples with code for how physical activity data can be processed, analyzed, and visualized in R.  
:::

## Getting Started

- Downloading R and RStudio <https://posit.co/download/rstudio-desktop/>
- Install the `devtools` package from CRAN  

```{r install devtools, echo = TRUE}
# The devtools package allows you to install packages from GitHub
if(!"devtools" %in% installed.packages()) install.packages("devtools")
```

- You can then install this presentation and the R packages listed in the description file

```{r install github}

if(!"devtools" %in% installed.packages(), devtools::install_github("bhelsel/CSACSM24"))

```





## Visualizations

- Adding text to see if the figure adjusts

![](images/comp_means.pdf)

